### 2020-03-24 Abstract/abstract.tex
Updated: Total proof-read and conclusion changed to approximate solution

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
In this study, a solution to delivering scalable real-time physics simulations is proposed. Although high performance computing simulations of physics related problems do exist, these are not real-time and do not model the real-time intricate interactions of rigid bodies for visual effect common in video games (favouring accuracy over real-time). As such, this study presents the first approach to real-time delivery of scalable, commercial grade, video game quality physics. We have termed this approach Aura Projection (AP). This approach takes the physics engine out of the player's machine and deploys it across standard cloud based infrastructures. The simulation world is divided into regions that are then allocated to multiple servers. A server maintains the physics for all simulated objects in its region. The contribution of this study is the ability to maintain a scalable simulation by allowing object interaction across region boundaries using predictive migration techniques. AP allows each object to project an aura that is used to determine object migration across servers to ensure seamless physics interactions between objects. AP allows player interaction at any point in real-time (influencing the simulation) in the same manner as any video game.

>>>>

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
In this study, a solution to delivering scalable real-time physics simulations is proposed. Although high performance computing simulations of physics related problems do exist, these are not real-time and do not model the real-time intricate interactions of rigid bodies for visual effect common in video games (favouring accuracy over real-time). As such, this study presents the first approach to real-time delivery of scalable, commercial grade, video game quality physics and is termed Aura Projection (AP). This approach takes the physics engine out of the player's machine and deploys it across standard cloud based infrastructures. The simulation world is divided into regions that are then allocated to multiple servers. A server maintains the physics for all simulated objects in its region. The contribution of this study is the ability to maintain a scalable simulation by allowing object interaction across region boundaries using predictive migration techniques. AP allows each object to project an aura that is used to determine object migration across servers to ensure seamless physics interactions between objects. AP allows player interaction at any point in real-time (influencing the simulation) in the same manner as any video game.

### 2020-03-24 ChapterBackground/chapterBackground.tex
Updated: Total proof-read and conclusion changed to approximate solution

The numerical integrator used to update the positions of objects in the system, uses a discrete time value. Each discrete time value is known as a time step. Each step simulates a set period of time. The shorter the time period simulated, the more physically accurate the simulation will be, however, each simulation step must take less time to execute than the time it is intending to simulate, otherwise the simulation would be unable to run in real-time. It is common for the time step to be a fixed value, though physics engines can support variable time steps.

@@

\subsection{Architectures}
Online game architectures fall into the following main categories client-server, peer-to-peer, hybrid peer-to-peer and distributed server (also referred to as multiserver). Each type has advantages and disadvantages and suitability depends on a number of factors, including number of players and whether or not the game is short-session-based or persistent.
%We discuss the wider topic of online gaming (client-hosted, P2P, server), challenges of latency, security/cheating and computational limits. Also some details of online gaming servers are discussed, specifically, typical server frame-times and network tick updates.

\subsubsection{Client-Server}
Client-Server architecture is a type of architecture in which the game is executed and game state is managed entirely by the server. Clients connect to the server and receive the necessary information about the game world. All messages and interactions are sent through the game server and the server is responsible for disseminating messages to the appropriate clients. The main challenge with this approach is scalability, as a single server can only support a limited number of players. This is normally solved with the addition of more servers, i.e. switching to a distributed server architecture. As all communication goes through the server, clients do not directly communicate with each other, meaning latency can be increased for interactions with other players.
As there is only one central version of the game state, consistency resolution is simplified to solving inconsistencies between server and client and the server treated as the correct master copy. In addition, fault-tolerance is simplified to keeping a single backup of the sate of the central server, although this adds more complexity and cost, and can reduce scalability. Client-Server architectures are less prone to cheating than peer-to-peer based architectures, as the game provider has complete control over the game state \cite{P2PForMMOs}.

@@

In the game world division method, there is a single instance of the game world and the workload is distributed either through each server being responsible for a region or a set of players. Division methods are discussed further in \ref{distrubted_virtual_env}. Even with the workload distributed between servers, one server may be become overloaded if the workload is too high, for example, by a large number of players in a region being managed by a single server. Dynamic load-balancing schemes can be employed to improve distribution between servers and prevent servers being overloaded when resources are available on other servers. The disadvantage of distributed servers is the need for complex hand-offs when players migrate or interact between servers and can lead to consistency problems between servers as latency will always exist between servers \cite{P2PForMMOs}.

@@

\subsection{Replica Manager}
RakNet provides a replica manager plugin, which supports the use of replicas and delta-encoding for the updates of their states. Replicas are replicated entities, hosted by one node and replicated on connected nodes. For example, a server is responsible for the updates if a set of entities, the entities are then replicated on all connected clients. This manager handles the broadcast of: existing game entities to new connections, new game entities to existing connections; deleted game entities to existing entities and updates of states of replicated entities \cite{RakNetReplica}.

@@

Maintaining consistency in the physics aspect of games when distributed is great challenge, this is due to the time-sensitive nature of physics (millisecond differences can result in completely divergent results, discussed further in \ref{divergingResults}) \cite{2012EnhancedDeadReckoning}.

@@

Although these correction techniques may be suitable for handling inconsistencies in simulation view between servers and clients, they are not suitable for maintaining consistency within a distributed simulation, which is the focus of this study. This is because in order to prevent diverging results, strong consistency of collisions must be maintained (discussed further in \ref{divergingResults}).

@@

Although there has been research utilising multiple servers to distribute the task of solving physics based problems (e.g., \cite{mashayekhi2018automatically}), to the best of our knowledge there is no literature describing real-time interactive physics exploiting the addition of servers to gain scalability. The closest work to our research is that carried out to seek scalability in terms of player numbers in online gaming in the field of Distributed Virtual Environments (DVEs).

@@

A microservice architecture for a game engine would create a more cost-efficient approach to cloud-hosted games and simulations, while also creating the potential for the scalability of individual systems on an on-demand basis. This study's problem is concentrated on separating out the physics system from the game engine and distributing it across multiple servers while maintaining the real-time streamed service to players. Through this distribution, this study aims to demonstrate that the physics aspect of a game engine can be horizontally scaled through the use of cloud computing.

@@

Streamed Gaming (also know as Gaming as a Service or Cloud Gaming) consists of cloud servers streaming to a player's device with player input being returned to the cloud server. The player's device acts as a thin client. The main benefit is that a player does not require expensive, powerful hardware, and games can be played on any operating system (e.g. Android, Linux and Mac). However, these benefits come at the cost of bandwidth and latency requirements \cite{iCloudAccess}.

The first successful demonstration of game streaming was in 2001 by G-cluster, whom publicly demonstrated game streaming over WiFi to a PDA in 2001 and launched a game-on-demand service in 2004 \cite{2016SurveyOnCloudGaming}. In the late 2000s, more companies introduced game streaming platforms such as OnLive \cite{OnLive}, Gaikai, and GameNow \cite{GameNow}. Due to financial difficulty in 2012, OnLive sold their patents to Sony and Gaikai and in 2012, Gaikai was bought by PSNOW \cite{PSNOW}.
Streamed Gaming services currently available include NVidia Geforce Now \cite{NVidiaGameStream}, PSNOW \cite{PSNOW}, Google Stadia\cite{Stadia} and Microsoft xCloud\cite{xCloud}. NVIDIA GRID hardware technology is targeted specifically at Streamed Gaming \cite{NVIDIACloud}.

@@

The reader has been provided a brief overview of the working of real-time physics and introduced to important terms and concepts. The current state of online gaming has been discussed, including the main challenges faced. In addition, aspects of online gaming that are relevant to this study have been presented in greater detail. The challenges in physics in online gaming are those of latency and consistency i.e. maintaining consistent states while latency exists between servers. Techniques for handling consistency on longer time scales (100s of milliseconds), such as interactions between players, already exist and compensation techniques (based on dead-reckoning) exist to reduce the perceived differences in physics state due high latency between clients and servers. However, neither of these are appropriate for distributing the simulation across multiple machines. Therefore this study intends to address this challenge through the use of a novel technique called AP.

>>>>

The numerical integrator used to update the positions of objects in the system, uses a discrete time value. Each discrete time value is known as a time step. Each step simulates a set period of time. The shorter the time period simulated, the more physically accurate the simulation will be. However, each simulation step must take less time to execute than the time it is intending to simulate, otherwise the simulation would be unable to run in real-time. It is common for the time step to be a fixed value, though physics engines can support variable time steps.

@@

\subsection{Architectures}
Online game architectures fall into the following main categories: client-server; peer-to-peer; hybrid peer-to-peer; and distributed server (also referred to as multiserver). Each type has advantages and disadvantages and suitability depends on a number of factors, including number of players and whether or not the game is short-session-based or persistent.
%We discuss the wider topic of online gaming (client-hosted, P2P, server), challenges of latency, security/cheating and computational limits. Also some details of online gaming servers are discussed, specifically, typical server frame-times and network tick updates.

\subsubsection{Client-Server}
Client-Server architecture is a type of architecture in which the game is executed and game state is managed entirely by the server. Clients connect to the server and receive the necessary information about the game world. All messages and interactions are sent through the game server and the server is responsible for disseminating messages to the appropriate clients. The main challenge with this approach is scalability, as a single server can only support a limited number of players \cite{Ferretti2018}. This is normally solved with the addition of more servers, i.e. switching to a distributed server architecture. As all communication goes through the server, clients do not directly communicate with each other, meaning latency can be increased for interactions with other players.
As there is only one central version of the game state, consistency resolution is simplified to solving inconsistencies between server and client and the server treated as the correct master copy. In addition, fault-tolerance is simplified to keeping a single backup of the sate of the central server, although this adds more complexity and cost, and can reduce scalability. Client-Server architectures are less prone to cheating than peer-to-peer based architectures, as the game provider has complete control over the game state \cite{P2PForMMOs}.

@@

In the game world division method, there is a single instance of the game world and the workload is distributed either through each server being responsible for a region or a set of players. Division methods are discussed further in Section \ref{distrubted_virtual_env}. Even with the workload distributed between servers, one server may be become overloaded if the workload is too high, for example, by a large number of players in a region being managed by a single server. Dynamic load-balancing schemes can be employed to improve distribution between servers and prevent servers being overloaded when resources are available on other servers. The disadvantage of distributed servers is the need for complex hand-offs when players migrate or interact between servers and can lead to consistency problems between servers as latency will always exist between servers \cite{P2PForMMOs}.

@@

\subsection{Replica Manager}
RakNet provides a replica manager plugin, which supports the use of replicas and delta-encoding for the updates of their states. Replicas are replicated entities, hosted by one node and replicated on connected nodes. For example, a server is responsible for the updates of a set of entities, the entities are then replicated on all connected clients. The replica manager handles the broadcast of: existing game entities to new connections, new game entities to existing connections; deleted game entities to existing entities and updates of states of replicated entities \cite{RakNetReplica}.

@@

Maintaining consistency in the physics aspect of games when distributed is great challenge. This is due to the time-sensitive nature of physics (millisecond differences can result in completely divergent results, discussed further in \ref{divergingResults}) \cite{2012EnhancedDeadReckoning}.

@@

Although these correction techniques may be suitable for handling inconsistencies in simulation view between servers and clients, they are not suitable for maintaining consistency within a distributed simulation, which is the focus of this research. This is because in order to prevent diverging results, strong consistency of collisions must be maintained (discussed further in \ref{divergingResults}).

@@

Although there has been research utilising multiple servers to distribute the task of solving physics based problems (e.g., \cite{mashayekhi2018automatically}), there is no known literature describing real-time interactive physics exploiting the addition of servers to gain scalability. The closest work to this challenge is that carried out to seek scalability in terms of player numbers in online gaming in the field of Distributed Virtual Environments (DVEs).

@@

A microservice architecture for a game engine would create a more cost-efficient approach to cloud-hosted games and simulations, while also creating the potential for the scalability of individual systems on an on-demand basis. This study concentrates on separating out the physics system from the game engine and distributing it across multiple servers while maintaining the real-time streamed service to players. Through this distribution, this study aims to demonstrate that the physics aspect of a game engine can be horizontally scaled through the use of cloud computing.

@@

Streamed Gaming (also know as Gaming as a Service or Cloud Gaming) consists of cloud servers streaming to a player's device with player input being returned to the cloud server. The player's device acts as a thin client. The main benefit is that a player does not require expensive, powerful hardware, and games can be played on any operating system (e.g. Android, Linux and Mac) \cite{D_Angelo_2015}. However, these benefits come at the cost of bandwidth and latency requirements \cite{iCloudAccess}.

The first successful demonstration of game streaming was in 2001 by G-cluster, who publicly demonstrated game streaming over WiFi to a PDA in 2001 and launched a game-on-demand service in 2004 \cite{2016SurveyOnCloudGaming}. In the late 2000s, more companies introduced game streaming platforms such as OnLive \cite{OnLive}, Gaikai, and GameNow \cite{GameNow}. Due to financial difficulty in 2012, OnLive sold their patents to Sony and Gaikai and in 2012, Gaikai was bought by PSNOW \cite{PSNOW}.
Streamed Gaming services currently available include NVidia Geforce Now \cite{NVidiaGameStream}, PSNOW \cite{PSNOW}, Google Stadia\cite{Stadia} and Microsoft xCloud\cite{xCloud}. NVIDIA GRID hardware technology is targeted specifically at Streamed Gaming \cite{NVIDIACloud}.

@@

The reader has been provided with a brief overview of the working of real-time physics and introduced to important terms and concepts. The current state of online gaming has been discussed, including the main challenges faced. In addition, aspects of online gaming that are relevant to this study have been presented in greater detail. The challenges in physics in online gaming are those of latency and consistency i.e. maintaining consistent states while latency exists between servers. Techniques for handling consistency on longer time scales (100s of milliseconds), such as interactions between players, already exist and compensation techniques (based on dead-reckoning) exist to reduce the perceived differences in physics state due high latency between clients and servers. However, neither of these are appropriate for distributing the simulation across multiple machines. Therefore this study intends to address this challenge through the use of a novel technique called AP.

### 2020-03-24 ChapterConclusion/chapterConclusion.tex
Updated: Total proof-read and conclusion changed to approximate solution

The correctness of collisions in AP have been measured and evaluated. The experiments conducted in this study establish that collisions remain correct in AP, even when objects interact while intersecting the boundary. This has been shown to be true while increasing each of the following factors: speed; latency and frame-time. Collisions remain correct up to the tolerance values for each factor and begin to show errors as soon as the tolerance is exceeded, demonstrating the correctness of the aura calculation used in AP. The holds true if thrashing collisions are excluded from the results. The number of thrashing collisions and amount of times thrashed were both measured, potential solutions to the thrashing problem were proposed in \ref{Thrashing} and it can be assumed that these errors will be fixed in future work.

@@

Acceleration could also be factored into the aura calculation. The velocity between time steps is often limited and using an acceleration tolerance would allow for smaller auras for slow moving objects. Smaller auras lead to better performance as it leads to fewer aura broadcasts and object migrations. Note that the acceleration tolerance could only apply to the displacement calculation of the local object, as the velocity of a potential remote object is unknown.
%Reference where smaller auras improving performance are explained

@@

An important feature of a physics engine is the ability to query the state of the simulation. A typical example of this would be a ray-cast, in which a ray is cast between two points in the simulation space to find intersecting bodies. %Ray-casts are commonly used in first-person shooting games, i.e. a ray is cast from the player's weapon to a given distance, and objects that intersect the ray are returned from the query.
Distributed physics adds an extra level of complexity to this as the query may concern spaces and/or objects being simulated on different servers. In the case of spatially-partitioned solutions, the query could overlap the region boundary and in order for the query to be completed, a round-trip of messages between the two servers would be required. If the query is blocking/synchronous (execution cannot continue until the query is complete), this would introduce a large amount of wait time (at least double the network latency). Another option would be to use asynchronous queries, the results of which get returned using a function callback as soon as the results have been received. However this is more complex than a query on a centralised simulation which can return the results synchronously without any network delay.

>>>>

The correctness of collisions in AP have been measured and evaluated. The experiments conducted in this study establish that AP provides at minimum, an approximate solution to maintaining correct collisions when the simulation is scaled, even when objects interact while intersecting the boundary. This has been shown to be true while increasing each of the following factors: speed; latency and frame-time. Collisions remain correct in the majority of cases up to the tolerance values for each factor and begin to show errors as soon as the maximum tolerance provided by the given aura size is exceeded, demonstrating the accuracy of the aura calculation used in AP. The holds true if thrashing collisions are excluded from the results. The number of thrashing collisions and amount of times thrashed were both measured, potential solutions to the thrashing problem were proposed in \ref{Thrashing} and it can be assumed that these errors will be fixed in future work.

@@

Acceleration could also be factored into the aura calculation. The velocity between time steps is often limited and using an acceleration tolerance would allow for smaller auras for slow moving objects. Smaller auras lead to better performance as it leads to fewer aura broadcasts and object migrations. However, the acceleration tolerance could only apply to the displacement calculation of the local object, as the velocity of a potential remote object is unknown.
%Reference where smaller auras improving performance are explained

@@

An important feature of a physics engine is the ability to query the state of the simulation. A typical example of this would be a ray-cast, in which a ray is cast between two points in the simulation space to find intersecting bodies. %Ray-casts are commonly used in first-person shooting games, i.e. a ray is cast from the player's weapon to a given distance, and objects that intersect the ray are returned from the query.
Distributed physics adds an extra level of complexity to this as the query may concern spaces and/or objects being simulated on different servers. In the case of spatially-partitioned solutions, the query could overlap the region boundary and in order for the query to be completed, a round-trip of messages between the two servers would be required. If the query is blocking/synchronous (execution cannot continue until the query is complete), this would introduce a large amount of wait time (at least double the network latency). Another option would be to use asynchronous queries, the results of which are returned using a function callback as soon as the results have been received. However this is more complex than a query on a centralised simulation which can return the results synchronously without any network delay.

### 2020-03-24 ChapterImplementation/chapterImplementation.tex
Updated: Total proof-read and conclusion changed to approximate solution

\chapter{Implementation}\label{Implementation}
The implementation of AP will be discussed in this section. The following work is based on the implementation of AP used in \cite{brown2019aura}, however, with many newer additions. We include full details of the implementation of AP to give a complete and whole solution.

@@

The limitation of the use of an overseer node is the number of clients that can connect is limited to the amount that can be handled by a single overseer node. In order to enable a scalable number of clients, multiple overseer nodes could be employed, dividing the workload between overseer nodes, allowing for many more clients to connect. This is illustrated in \ref{MultiOverseer}.

@@

Overseer nodes could be located at the edge, geographically closer to the clients, as illustrated in \ref{EdgeOverseer}. The advantage of this approach is, there is a reduced latency between clients and overseers, meaning lower delay times between clients and the workload being carried out by the overseers, such as game logic. The disadvantage of this approach is that there is higher latency between overseer nodes, which are located far apart, geographically.

@@

In the current system architecture, the rendering work is carried out on the client machine. An alternative to this is to have the rendering performed on dedicated render nodes in the cloud, as illustrated in \ref{RenderNodes}. The results video could then be streamed to clients. This has the advantage of enabling low computational power devices, such as mobile devices, to display higher quality graphics, beyond what would be achievable on low computational power devices. Additionally, it means there is very low latency between the physics simulation and the renderer, meaning what is displayed is closer to the actual current state of the physics simulation. The disadvantage of this approach is the latency time between the client and the renderer.

@@

When a server begins communicating an aura, a message is added to the aura creation/update buffer. This message consists of the position of the centre of the aura and the radius of the aura. % and a flag signifying whether or not this is a creation message or not. Flag is not currently used as often causes issue in the corner case, possibly due to multiple types of messages concerncing one object being sent and recieved in the same tick.
The contents of the buffer are then broadcast to all neighbouring servers. When aura creation/update messages are received and an aura for that object does not already exist, a new aura, a sphere trigger volume, is created at that position with the given radius.
Whenever an object projecting an aura moves, and update message is added to the aura creation/update buffer, containing the new position and radius of the aura (aura radius is currently fixed, but may become dynamic in the future). The contents of the buffer are broadcast to all neighbouring servers. When the message is received, the positions of the auras are updated. 
If an object leaves aura trigger boundary, either by moving away from the boundary or by being migrated to another server, a delete aura message is added to the aura delete buffer. The contents of the buffer are then broadcast to all neighbouring servers. When the message is received, the aura trigger volume is removed from the simulation.

@@

\subsection{Client-created static objects}\label{Client-CreatedReplicas}
In addition to replica objects, static objects can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers, this could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low. They do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects, this avoids the need for extra communication when server boundaries change. With these trade-offs in mind, the decision for AP was made to create all static objects on all servers, when static objects are used in a simulation.

@@

\subsection{Future work}
Currently when server created objects are migrated between servers, the client destroys the replica and a new replica is created. This is because the simulated object is destroyed on one server and re-created on another. A more efficient way of dealing with this, would be a procedure for handing over server-ownership of replicas, removing the need for replicas to be destroyed and recreated. This would reduce processing overhead on the client, as there is no longer a need to destroy and recreate objects when server migrations take place. In addition, this would also prevent any chance of objects appearing to `flicker' due to any delays between being destroyed and recreated.

>>>>

\chapter{Implementation}\label{Implementation}
The implementation of AP will be discussed in this section. The following work is based on the implementation of AP used in \cite{brown2019aura}, however, with many newer additions. Full details of the implementation of AP have been included to give a complete and whole solution.

@@

The limitation of the use of an overseer node is the number of clients that can connect is limited to the amount that can be handled by a single overseer node. In order to enable a scalable number of clients, multiple overseer nodes could be employed, dividing the workload between overseer nodes, allowing for many more clients to connect. This is illustrated in Fig. \ref{MultiOverseer}.

@@

Overseer nodes could be located at the edge, geographically closer to the clients, as illustrated in Fig. \ref{EdgeOverseer}. The advantage of this approach is, there is a reduced latency between clients and overseers, meaning lower delay times between clients and the workload being carried out by the overseers, such as game logic. The disadvantage of this approach is that there is higher latency between overseer nodes, which are located far apart, geographically.

@@

In the current system architecture, the rendering work is carried out on the client machine. An alternative to this is to have the rendering performed on dedicated render nodes in the cloud, as illustrated in Fig. \ref{RenderNodes}. The results video could then be streamed to clients. This has the advantage of enabling low computational power devices, such as mobile devices, to display higher quality graphics, beyond what would be achievable on low computational power devices. Additionally, it means there is very low latency between the physics simulation and the renderer, meaning that what is displayed is closer to the actual current state of the physics simulation. The disadvantage of this approach is the latency time between the client and the renderer.

@@

When a server begins communicating an aura, a message is added to the aura creation/update buffer. This message consists of the position of the centre of the aura and the radius of the aura. % and a flag signifying whether or not this is a creation message or not. Flag is not currently used as often causes issue in the corner case, possibly due to multiple types of messages concerncing one object being sent and recieved in the same tick.
The contents of the buffer are then broadcast to all neighbouring servers. When aura creation/update messages are received and an aura for that object does not already exist, a new aura, a sphere trigger volume, is created at that position with the given radius.
Whenever an object projecting an aura moves, an update message is added to the aura creation/update buffer, containing the new position and radius of the aura (aura radius is currently fixed, but may become dynamic in the future). The contents of the buffer are broadcast to all neighbouring servers. When the message is received, the positions of the auras are updated. 
If an object leaves aura trigger boundary, either by moving away from the boundary or by being migrated to another server, a delete aura message is added to the aura delete buffer. The contents of the buffer are then broadcast to all neighbouring servers. When the message is received, the aura trigger volume is removed from the simulation.

@@

\subsection{Client-created static objects}\label{Client-CreatedReplicas}
In addition to replica objects, static objects can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers. This could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low. They do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects. This avoids the need for extra communication when server boundaries change. With these trade-offs in mind, the decision for AP was made to create all static objects on all servers, when static objects are used in a simulation.

@@

\subsection{Future work}
Currently, when server created objects are migrated between servers, the client destroys the replica and a new replica is created. This is because the simulated object is destroyed on one server and re-created on another. A more efficient way of dealing with this, would be a procedure for handing over server-ownership of replicas, removing the need for replicas to be destroyed and recreated. This would reduce processing overhead on the client, as there is no longer a need to destroy and recreate objects when server migrations take place. In addition, this would also prevent any chance of objects appearing to `flicker' due to any delays between being destroyed and recreated.

### 2020-03-24 ChapterIntroduction/chapterIntroduction.tex
Updated: Total proof-read and conclusion changed to approximate solution

Traditionally real-time simulations are confined to a single machine, including real-time physics. 
Recent cloud technology has made available, lower cost, low-latency and high-performance machines, including GPU-enabled instances. This means, real-time simulations can now run in the cloud as well as being economically viable. It is possible to rapidly provision resources on demand, allowing for fast scaling up and down of resources as needed.

@@

Server authoritative physics would also allow over the network multiplayer games to rely heavily on physical elements. In games, typically only a limited part of the game is simulated on the server side and not physics, with focus on player-player interaction and not player-environment (physics) interaction, as this is a limit of existing technology. 
Synchronising physics states between players and servers is not possible due to latency, two players will have different views of the physical state of the world, based on their latency, interactions from those player's with the physical state will cause divergence, which may not be possible to converge. This is normally worked around by not having physics affect gameplay and using kinematic sequences, ensuring determinism. A single physics simulation does not encounter these problems of multiple physics states, which have to be synchronised and can lead to divergent physics states.

@@

\section{Streamed Gaming}
Streamed gaming is an active area of research (e.g., \cite{GARCIAVALLS}), where the game is executed on the server-side, the graphical output is streamed to a remote player and input is sent from the player to the server-side.
Streamed gaming reduces the computing power requirements on player hardware, even allowing gaming on mobile devices with graphical quality beyond what would be possible if the game were running entirely on the device \cite{CloudGamingArchPerf}. In addition, this removes the need for players to continuously upgrade their hardware \cite{LatencyoOfCloudGamingSystems}.

@@

\section{Summary}
Real-time physics is an integral part of games technology. There has been recent commercial interest in using cloud computing to scale up games, including the most challenging aspect, the physics simulation component. In addition, scaling up physics simulations provides an increase in possible size and complexity of virtual worlds used in online games (a greater number of simulated entities and an increase in systems that are expensive to simulate). If achieved, real-time physics engines can be scaled beyond what is currently possible on a player's machine or single server and provides elastic scaling of resources for use in streamed gaming. Despite the recent interest and advantages, no academic literature has been produced addressing scalable real-time physics.

>>>>

Traditionally real-time simulations are confined to a single machine, including real-time physics. 
Recent cloud technology has made available lower cost, low-latency and high-performance machines, including GPU-enabled instances. This means, real-time simulations can now run in the cloud as well as being economically viable. It is possible to rapidly provision resources on demand, allowing for fast scaling up and down of resources as needed.

@@

Server authoritative physics would also allow over the network multiplayer games to rely heavily on physical elements. In games, typically only a limited part of the game is simulated on the server side and not physics, with focus on player-player interaction and not player-environment (physics) interaction, as this is a limit of existing technology. 
Synchronising physics states between players and servers is not possible due to latency, two players will have different views of the physical state of the world, based on their latency, interactions from those players with the physical state will cause divergence, which may not be possible to converge. This is normally worked around by not having physics affect gameplay and using kinematic sequences, ensuring determinism. A single physics simulation does not encounter these problems of multiple physics states, which have to be synchronised and can lead to divergent physics states.

@@

\section{Streamed Gaming}
Streamed gaming is an active area of research (e.g., \cite{GARCIAVALLS}), where the game is executed on the server-side, the graphical output is streamed to a remote player and input is sent from the player to the server-side.
Streamed gaming reduces the computing power requirements on player hardware, even allowing gaming on mobile devices with graphical quality beyond what would be possible if the game was running entirely on the device \cite{CloudGamingArchPerf}. In addition, this removes the need for players to continuously upgrade their hardware \cite{LatencyoOfCloudGamingSystems}.

@@

\section{Summary}
Real-time physics is an integral part of games technology. There has been recent commercial interest in using cloud computing to scale up games, including the most challenging aspect, the physics simulation component. In addition, scaling up physics simulations provides an increase in possible size and complexity of virtual worlds used in online games (a greater number of simulated entities and an increase in systems that are expensive to simulate). If achieved, real-time physics engines can be scaled beyond what is currently possible on a player's machine or single server and provides elastic scaling of resources for use in streamed gaming. Despite the recent interest and advantages, no academic literature (with the exception of \cite{brown2019aura} based on the work of this study) addressing scalable real-time physics has been produced, to date.
